Comparing fluorometric methods ( in vivo vs extracted) for cyanoHAB monitoring in six Rhode Island ponds

### 2024-02-20 manuscript/manuscript_old.qmd
made edits but have no idea what i did

Within 24 hours of collection, three replicates (i.e. lab replicates) from each field sample was analyzed using the PhycoProbe, CyanoFluor, and the Trilogy (in-vivo), as well as filtered onto 0.7 µm pre-ashed glass fiber filters and frozen at -20 °C for solvent-based extraction. Sample water filtered through a 0.22 µm syringe filter was used to correct for yellow substances on the PhycoProbe, CyanoFluor, and Trilogy in vivo module. Chlorophyll *a* extraction using 90% acetone followed after a 20 minute period in a sonicating water bath. Phycocyanin extraction followed a 15 minute period in a sonicating water bath and used a 50 mM phosphate buffer. Extracted samples were analyzed using a Turner Trilogy with a chla na module (chlorophyll) and an orange module (phycocyanin)(\@kasinak2015benchtop; \@welschmeyer1994fluorometric).

@@

!!!REWRITE ME - Changed from biov to cells/ml, only one fig now!!!!

Cyanophytes were the dominant taxa by relative biovolume at Curran (0.75), Mashapaug (0.87), and Warwick (0.89); Chrysophytes at Barber (0.39) and Yawgoo (0.44); and Bacillariophytes at Indian (0.41)(@fig-cellsml).  Total biovolume concentrations were highest at Mashapaug (2.12x10^7^), Warwick (1.44x10^7^), and Curran (3.22x10^6^)(@fig-bv). Microcystis was the dominant genus at Mashapaug and Warwick, while Chrysosporum was dominant at Curran (@tbl-supp).

![Compare Cells/mL of divisions by waterbody](images/fig4_total_cellsml_bar.jpg){#fig-cellsml width="8.5in" height="5.25in"}

@@

![Compare fluorometer chlorophyll to cyanobacteria cell counts](images/fig5_chl_vs_cells.jpg){#fig-chlcount width="8.5in" height="5.25in"}

![Compare fluorometer phycocyanin to cyanobacteria cell counts](images/fig6_phyco_vs_cells.jpg){#fig-phycocount width="8.5in" height="5.25in"}

>>>>

Within 24 hours of collection, three replicates (i.e. lab replicates) from each field sample was analyzed using the PhycoProbe, CyanoFluor, and the Trilogy (in-vivo), as well as filtered onto 0.7 µm pre-ashed glass fiber filters and frozen at -20 °C for solvent-based extraction. Sample water filtered through a 0.22 µm syringe filter was used to correct for yellow substances on the PhycoProbe, CyanoFluor, and Trilogy in vivo module. Chlorophyll *a* extraction using 90% acetone followed after a 20 minute period in a sonicating water bath. Phycocyanin extraction followed a 15 minute period in a sonicating water bath and used a 50 mM phosphate buffer. Extracted samples were analyzed using a Turner Trilogy with a chla na module (chlorophyll) and an orange module (phycocyanin)(@kasinak2015benchtop; @welschmeyer1994fluorometric).

@@

Cyanophytes were the dominant taxa by relative biovolume at Curran (0.75), Mashapaug (0.87), and Warwick (0.89); Chrysophytes at Barber (0.39) and Yawgoo (0.44); and Bacillariophytes at Indian (0.41)(@fig-relbv).  Total biovolume concentrations were highest at Mashapaug (2.12x10^7^), Warwick (1.44x10^7^), and Curran (3.22x10^6^)(@fig-bv). Microcystis was the dominant genus at Mashapaug and Warwick, while Chrysosporum was dominant at Curran (@tbl-supp).

![Compare Relative Biovolume of divisions by waterbody](images/fig4_rel_bio_bar.jpg){#fig-relbv width="8.5in" height="5.25in"}

![Compare Total Biovlume of divisions by waterbody](images/fig5_total_bio_bar.jpg){#fig-bv width="8.5in" height="5.25in"}

@@

![Compare fluorometer chlorophyll to cyanobacteria cell counts](images/fig6_chl_vs_cells.jpg){#fig-chlcount width="8.5in" height="5.25in"}

![Compare fluorometer phycocyanin to cyanobacteria cell counts](images/fig7_phyco_vs_cells.jpg){#fig-phycocount width="8.5in" height="5.25in"}
